---
title: "GM-UAW Cohort Study"
subtitle: "Prostate cancer incidence"
date: \today
# fontsize: 10pt
output:
  beamer_presentation:
    # latex_engine: lualatex
    includes:
      in_header: "~/HeadRs/BeamerHead.sty"
  keep_tex: true
---

```{r setup, include=F}
knitr::opts_chunk$set(echo = F,
											warning = F,
											message = F,
											cache = F,
											fig.align = "center",
											fig.pos = "H",
											results = "asis")

library(here); library(tidyverse); library(data.table);
library(lubridate); library(xtable)

source("~/HeadRs/00-my-theme.R")
incidence.key <- fread(here::here("cancer incidence", "cancer-key-expanded.tsv"))

```

# Four analyses

\linespread{1.25}

1. Men still alive in 1973
1. Men still alive in 1985, excluding the 134 men who had prostate cancer before 1985
1. Men still alive in 1985, excluding the 134, weighted by probability of not dying of natural causes
1. Men still alive in 1985, excluding the 134, weighted by stabilized inverse probability with truncation at the 99^th^ percentile of weights ($sw_i$ = 330)

Follow-up ends on the date of death, prostate cancer incidence, LTFU (outcome-free at age 108.39 years), or 2015-12-31, whichever comes first.

All analyses restricted to plants 1 and 2

Summary of population charactersitcs in the slides to follow.

#

\renewcommand\arraystretch{1.1}

\begin{table}[H]
\begin{adjustbox}{scale=0.75}
\centering
```{r}
tab1.1 <- readRDS(file = paste0(
					here::here('left truncation/resources/'),
					"popchar1.tab1.rds"
				))

n.which <- which(grepl("\\$n", tab1.1[,"spread.which"]))
med.which <- which(grepl("median", tab1.1[,"spread.which"]) &
									 	!is.na(tab1.1[,3]))

tab1.1[med.which, 2] <- paste0("$(", gsub("\\$", "", tab1.1[med.which, 2]),
															",\\ ",
															gsub("\\$", "", tab1.1[med.which, 3]),
															")$")

tab1.1 <- cbind(rownames(tab1.1), tab1.1[,1:2])

colnames(tab1.1)[1] <- "\\multicolumn{3}{l}{Men still alive in 1973}\\\\%"

print.xtable(xtable(tab1.1[-c(6:8, 12:16),],
										align = "llrl"))

```
\end{adjustbox}
\end{table}

#

\begin{table}[H]
\begin{adjustbox}{scale=0.75}
\centering
```{r}
tab1.2 <- readRDS(file = paste0(
					here::here('left truncation/resources/'),
					"popchar2.tab1.rds"
				))

n.which <- which(grepl("\\$n", tab1.2[,"spread.which"]))
med.which <- which(grepl("median", tab1.2[,"spread.which"]) &
									 	!is.na(tab1.2[,3]))

tab1.2[med.which, 2] <- paste0("$(", gsub("\\$", "", tab1.2[med.which, 2]),
															",\\ ",
															gsub("\\$", "", tab1.2[med.which, 3]),
															")$")

tab1.2 <- cbind(rownames(tab1.2), tab1.2[,1:2])

colnames(tab1.2)[1] <- "\\multicolumn{3}{l}{Men still alive in 1985}\\\\%"

print.xtable(xtable(tab1.2[-c(6:8, 12:16),],
										align = "llrl"))

```
\end{adjustbox}
\end{table}

#

\begin{table}[H]
\begin{adjustbox}{scale=0.75}
\centering
```{r}
tab1.3 <- readRDS(file = paste0(
					here::here('left truncation/resources/'),
					"popchar3.tab1.rds"
				))

n.which <- which(grepl("\\$n", tab1.3[,"spread.which"]))
med.which <- which(grepl("median", tab1.3[,"spread.which"]) &
									 	!is.na(tab1.3[,3]))

tab1.3[med.which, 2] <- paste0("$(", gsub("\\$", "", tab1.3[med.which, 2]),
															",\\ ",
															gsub("\\$", "", tab1.3[med.which, 3]),
															")$")

tab1.3 <- cbind(rownames(tab1.3), tab1.3[,1:2])

colnames(tab1.3)[1] <- "\\multicolumn{3}{l}{Men still alive in 1985}\\\\%"

print.xtable(xtable(tab1.3[-c(6:8, 13:16),],
										align = "llrl"))

```
\end{adjustbox}
\end{table}

# Stabilized weight

\linespread{1.25}

$$sw_i = \frac{1 - \bar p}{1 - \hat p_i}$$
where $\hat p_i$ is the $i$^th^ individual's predicted probability of surviving to year 1985, and $\bar p$ is the overall survival proportion i.e. $\frac{1}{n}\sum^n_i\hat p_i$.

# Cox PH Models

\linespread{1.25}

- Cumulative exposure to straight, soluble, and synthetic MWFs (lagged 20 years)
- Splined calendar year
- Splined year of hire
- Race
- Plant
- Risk sets indexed by age

#

## 1. Men still alive in 1973 \vspace{2.5pt}

\begin{table}[H]
\begin{adjustbox}{scale=0.75}
\centering
```{r}
coef1.tab <- readRDS(here::here('reports/left truncation/prostate cancer/resources/', "pr_mod1.tab.rds"))

colnames(coef1.tab)[5] <- "(95\\% CI)"
coef1.tab <- coef1.tab[,-c("events", "p")]
coef1.tab[,"n"] <- as.character(round(unlist(coef1.tab[,"n"]), 0))
coef1.tab[,"Covariate"] <- gsub(" 5$", "", unlist(coef1.tab[,"Covariate"]))

print.xtable(xtable(coef1.tab))

```
\end{adjustbox}
\end{table}

#

## 2. Men still alive in 1985 \vspace{2.5pt}

\begin{table}[H]
\begin{adjustbox}{scale=0.75}
\centering
```{r}
coef2.tab <- readRDS(here::here('reports/left truncation/prostate cancer/resources/', "pr_mod2.tab.rds"))

colnames(coef2.tab)[5] <- "(95\\% CI)"
coef2.tab <- coef2.tab[,-c("events", "p")]
coef2.tab[,"n"] <- as.character(round(unlist(coef2.tab[,"n"]), 0))
coef2.tab[,"Covariate"] <- gsub(" 5$", "", unlist(coef2.tab[,"Covariate"]))

print.xtable(xtable(coef2.tab))

```
\end{adjustbox}
\end{table}

# 

## 3. Men still alive in 1985, weighted by probability of not being dead due to natural causes in 1985 \vspace{2.5pt}

\begin{table}[H]
\begin{adjustbox}{scale=0.75}
\centering
```{r}
coef3.tab <- readRDS(here::here('reports/left truncation/prostate cancer/resources/', "pr_mod3.tab.rds"))

colnames(coef3.tab)[5] <- "(95\\% CI)"
coef3.tab <- coef3.tab[,-c("events", "p")]
coef3.tab[,"n"] <- as.character(round(unlist(coef3.tab[,"n"]), 0))
coef3.tab[,"Covariate"] <- gsub(" 5$", "", unlist(coef3.tab[,"Covariate"]))

print.xtable(xtable(coef3.tab))

```
\end{adjustbox}
\end{table}

#

## 4. Men still alive in 1985, weighted by the stabilized inverse probability with truncation \vspace{2.5pt}

\begin{table}[H]
\begin{adjustbox}{scale=0.75}
\centering
```{r}
coef4.tab <- readRDS(here::here('reports/left truncation/prostate cancer/resources/', "pr_mod4.tab.rds"))

colnames(coef4.tab)[5] <- "(95\\% CI)"
coef4.tab <- coef4.tab[,-c("events", "p")]
coef4.tab[,"n"] <- as.character(round(unlist(coef4.tab[,"n"]), 0))
coef4.tab[,"Covariate"] <- gsub(" 5$", "", unlist(coef4.tab[,"Covariate"]))

print.xtable(xtable(coef4.tab))

```
\end{adjustbox}
\end{table}


# Results by MWF type

\begin{table}[H]
\begin{adjustbox}{scale=0.7}
\centering
```{r}
str.tab <- cbind(
	coef1.tab[grep("str", zoo::na.locf(coef1.tab[,Covariate])), 2],
	NA,
	coef1.tab[grep("str", zoo::na.locf(coef1.tab[,Covariate])),-c(1, 2, 5)],
	NA,
	coef2.tab[grep("str", zoo::na.locf(coef2.tab[,Covariate])), c(3, 4)],
	NA,
	coef3.tab[grep("str", zoo::na.locf(coef3.tab[,Covariate])), c(3, 4)],
	NA,
	coef4.tab[grep("str", zoo::na.locf(coef4.tab[,Covariate])), c(3, 4)])

colnames(str.tab) <- sapply(1:ncol(str.tab), function (x) {paste(rep("%", x), collapse = "")})

print.xtable(xtable(str.tab,
										align = "llcccccccccccc"),
						 add.to.row = list(
						 	pos = list(0),
						 	command = paste0(
						 		"Cumulative straight & &",
						 		"\\multicolumn{2}{c}{1973} & &",
						 		"\\multicolumn{2}{c}{1985} & &",
						 		"\\multicolumn{2}{c}{1985, weighted} & &",
						 		"\\multicolumn{2}{c}{1985, s. weight}",
						 		"\\\\\n",
						 		"\\cline{3-4}",
						 		"\\cline{6-7}",
						 		"\\cline{9-10}",
						 		"\\cline{12-13}",
						 		"\n",
						 		paste0(rep("&& n & HR", 4), collapse = ""),
						 		"\\\\")
						 ))

```
\end{adjustbox}
\end{table}

\begin{table}[H]
\begin{adjustbox}{scale=0.7}
\centering
```{r}
sol.tab <- cbind(
	coef1.tab[grep("sol", zoo::na.locf(coef1.tab[,Covariate])), 2],
	NA,
	coef1.tab[grep("sol", zoo::na.locf(coef1.tab[,Covariate])),-c(1, 2, 5)],
	NA,
	coef2.tab[grep("sol", zoo::na.locf(coef2.tab[,Covariate])), c(3, 4)],
	NA,
	coef3.tab[grep("sol", zoo::na.locf(coef3.tab[,Covariate])), c(3, 4)],
	NA,
	coef4.tab[grep("sol", zoo::na.locf(coef4.tab[,Covariate])), c(3, 4)])

colnames(sol.tab) <- sapply(1:ncol(sol.tab), function (x) {paste(rep("%", x), collapse = "")})

print.xtable(xtable(sol.tab,
										align = "llcccccccccccc"),
						 add.to.row = list(
						 	pos = list(0),
						 	command = paste0(
						 		"Cumulative soluble & &",
						 		"\\multicolumn{2}{c}{1973} & &",
						 		"\\multicolumn{2}{c}{1985} & &",
						 		"\\multicolumn{2}{c}{1985, weighted} & &",
						 		"\\multicolumn{2}{c}{1985, s. weight}",
						 		"\\\\\n",
						 		"\\cline{3-4}",
						 		"\\cline{6-7}",
						 		"\\cline{9-10}",
						 		"\\cline{12-13}",
						 		"\n",
						 		paste0(rep("&& n & HR", 4), collapse = ""),
						 		"\\\\")
						 ))

```
\end{adjustbox}
\end{table}

\begin{table}[H]
\begin{adjustbox}{scale=0.7}
\centering
```{r}
syn.tab <- cbind(
	coef1.tab[grep("syn", zoo::na.locf(coef1.tab[,Covariate])), 2],
	NA,
	coef1.tab[grep("syn", zoo::na.locf(coef1.tab[,Covariate])),-c(1, 2, 5)],
	NA,
	coef2.tab[grep("syn", zoo::na.locf(coef2.tab[,Covariate])), c(3, 4)],
	NA,
	coef3.tab[grep("syn", zoo::na.locf(coef3.tab[,Covariate])), c(3, 4)],
	NA,
	coef4.tab[grep("syn", zoo::na.locf(coef4.tab[,Covariate])), c(3, 4)])

colnames(syn.tab) <- sapply(1:ncol(syn.tab), function (x) {paste(rep("%", x), collapse = "")})

print.xtable(xtable(syn.tab,
										align = "llcccccccccccc"),
						 add.to.row = list(
						 	pos = list(0),
						 	command = paste0(
						 		"Cumulative synthetic & &",
						 		"\\multicolumn{2}{c}{1973} & &",
						 		"\\multicolumn{2}{c}{1985} & &",
						 		"\\multicolumn{2}{c}{1985, weighted} & &",
						 		"\\multicolumn{2}{c}{1985, s. weight}",
						 		"\\\\\n",
						 		"\\cline{3-4}",
						 		"\\cline{6-7}",
						 		"\\cline{9-10}",
						 		"\\cline{12-13}",
						 		"\n",
						 		paste0(rep("&& n & HR", 4), collapse = ""),
						 		"\\\\")
						 ))

```
\end{adjustbox}
\end{table}