---
title: "Risk of incident prostate cancer risk, weighted by survival to 1985"
subtitle: "GM-UAW Cohort Study"
date: \today
fontsize: 11pt
format:
  docx:
    reference-doc: '/Users/kevinchen/HeadRs/rmd_templates/style-guide-arial-11.docx'
execute:
  freeze: auto
editor_options: 
  chunk_output_type: inline
bibliography: "`r path.expand('~/Box Sync/papers/EnvEpi.bib')`"
csl: "`r path.expand('~/HeadRs/csl/AMA.csl')`"
---

\renewcommand\arraystretch{1.1}
\onehalfspacing

```{r setup, include=F}
knitr::opts_chunk$set(echo = F,
											warning = F,
											message = F,
											cache = F,
											fig.align = "center",
											fig.pos = "H",
											results = "asis")

library(here); library(data.table);
library(flextable)

source("~/HeadRs/00-my-theme.R")
incidence.key <- fread(here::here("cancer incidence", "cancer-key-expanded.tsv"))

sanitize.tex <- function(j, sanitize = T, firstcol = T) {
	j <- as.data.table(as.data.frame(j))
	col.names <- names(j)
	col.names <- col.names[if (firstcol) {
		1:length(col.names)} else {-1}]
	j[,(col.names):=(
		lapply(col.names, function(x) {
			x <- get(x)
			if (sanitize) {
				x <- gsub("\\\\,|\\\\ ", " ", x)
				x <- gsub("\\$", "", x)
				x
			} else {x}
		}))]
	j
}
```

# The problem

The Cohort includes all persons who have worked at least 3 years and were hired in the years 1938 through 1981. However, follow-up for cancer incidence (Michigan Cancer Registry) does not begin until 1985. Not all who entered the cohort prior to 1985 were still at-risk of the outcome in 1985. Loss to follow-up may have been differential by underlying health. In particular, healthier individuals may have had lower exposure and a greater probability of surviving to 1985 outcome-free. In this case, the resulting population at-risk in 1985 would be the result of the so-called healthy worker survivor bias.[@Hernan_2004] (There may also be those who have already experienced incident cancer prior to 1985, but were not recorded as such, but this outcome misclassification problem is not the focus of this analysis.)

```{sh, eval=F}
lualatex -jobname=dag <<theend
\documentclass[11pt, border={5pt 5pt 5pt 5pt}]{standalone}
\usepackage[no-math]{fontspec}
\setmainfont{Arial}
\usepackage[italic]{mathastext}
\input{/Users/kevinchen/HeadRs/StatHead.sty}

\usepackage[active,tightpage,psfixbb]{preview}
\PreviewEnvironment{tikzpicture}
\setlength\PreviewBorder{0pt}

\begin{document}
  \begin{tikzpicture}[
      >= stealth, auto,
      node distance = 2.5cm, semithick, scale=0.85,
      thick, fill=white]
  \tikzstyle{every state}=[
      shape = rectangle, align = center, draw = none,
      inner sep=2pt, outer sep=2pt,
      minimum size=20pt]
  \node[state] (Ustar) {\$U^*\$};
  \node[state] (L) [below of=Ustar] {\$L\$};
  \node[state] (U) [below of=L]{\$U\$};
  \node[state] (E) [right of=L]{\$E\$};
  \node[state] (C) [right of=E, draw=black]{\$C\$};
  \node[state] (Y) [right of=C]{\$Y\$};
  \path[->] (Ustar) edge (E);
  \path[->] (Ustar) edge (C);
  \path[->] (U) edge (L);
  \path[->] (U) edge [bend right=15] (Y);
  \path[->] (L) edge [bend right=35] (C);
  \end{tikzpicture}
\end{document}
theend
rm dag.aux dag.out
```

![Adapted from Figure 6c in Hernán et al. (2004). Take $L$, $E$, $C$, and $Y$ to represent underlying health, occupational exposure, censoring, and cancer incidence, respectively.](dag.pdf)


# Workaround

We have no measure of underlying health $L$, but we proceeded by modeling the censoring mechanism $\mathbb P_{i}(C_t \mid X_t ,C_{t - 1} = 0)$ for individual $i$ at year $t$, conditional on observed covariates $X_t$ that may serve as instruments for unmeasured $L$. We then used the modeled censoring probability $g_i(C_t\mid X_t ,C_{t - 1} = 0)$ to construct stabilized weights $sw_i$ for each individual in $t = 1985$ where
$$sw_i = \frac{1 - \frac{1}{n}\sum^n_i \mathbb \prod_t^{1984} g_i(C_{t}  = 0\mid X_{t} ,C_{(t - 1)} = 0)}{1 - \prod_t^{1984} g_i(C_{t}  = 0 \mid X_{t} ,C_{(t - 1)} = 0)}.$$

The parameters of model $g_i(C_t \mid X_t, C_{t - 1} = 0)$ were estimated using a pooled logistic regression for the log-odds of dying due to natural causes by the end of each year of follow-up $t = \{1, \dots, T\}$, conditional on the $P$-length covariate vector $X_t = x_t$ and uncensored status at the begninnig of that person-year $C_{t - 1} = 0$.

$$\log\frac{\widehat{\mathbb P}\left(C_{t} = 1 \mid X_{t} = x_t, \hat{\beta}, C_{t - 1} = 0\right)}{1 - \widehat{\mathbb P}\left(C_{t} = 1 \mid X_{t} = x_t, \hat{\beta}, C_{t - 1} = 0\right)}
= \hat{\beta}_{0} + \hat{\beta}_1 X_{1t} +  \cdots + \hat{\beta}_P X_{Pt}.$$

The covariates $X_t$ included:

- Years since hire (quartiles or splined)
- Age (quartiles or splined)
- Plant
- Race (black or white)
- Sex
- Proportion of year spent in assembly, machining (includes grinding), and off (quartiles)
- Cumulative time spent off (quartiles)
- Year of hire (quartiles)
- Cumulative exposure to straight, soluble, and synthetic MWFs (quartiles)
- Employment status

Further information on the estimation of $g$ can be found on [Box](https://berkeley.app.box.com/folder/114285681411).

# Four analyses

1. Men still alive in 1973
1. Men still alive in 1985, excluding the 134 men who had prostate cancer before 1985
1. Men still alive in 1985, excluding the 134, weighted by probability of not dying of natural causes
1. Men still alive in 1985, excluding the 134, weighted by stabilized inverse probability with truncation at the 99^th^ percentile of weights ($sw_i$ = 330)

Follow-up ends on the date of death, prostate cancer incidence, LTFU (outcome-free at age 108.39 years), or 2015-12-31, whichever comes first.

All analyses restricted to plants 1 and 2

Summary of population characteristics in the slides to follow.

#

\begin{table}[H]
\centering
```{r}
tab1.1 <- readRDS(here::here('resources/', 'popchar1.tab1.rds'))

n.which <- which(grepl("\\$n", tab1.1[,"spread.which"]))
med.which <- which(grepl("median", tab1.1[,"spread.which"]) &
									 	!is.na(tab1.1[,3]))

tab1.1[med.which, 2] <- paste0("$(", gsub("\\$", "", tab1.1[med.which, 2]),
															",\\ ",
															gsub("\\$", "", tab1.1[med.which, 3]),
															")$")

tab1.1 <- cbind(rownames(tab1.1), tab1.1[,1:2])

colnames(tab1.1)[1] <- "\\multicolumn{3}{l}{Men still alive in 1973}\\\\%"

tab1.1[-c(6:8, 12:16),] %>%
	sanitize.tex(firstcol = F) %>% xtable(align = "llrl") %>% print.xtable

```
\end{table}

#

\begin{table}[H]
\centering
```{r}
tab1.2 <- readRDS(file = here::here('resources/', 'popchar2.tab1.rds'))

n.which <- which(grepl("\\$n", tab1.2[,"spread.which"]))
med.which <- which(grepl("median", tab1.2[,"spread.which"]) &
									 	!is.na(tab1.2[,3]))

tab1.2[med.which, 2] <- paste0("$(", gsub("\\$", "", tab1.2[med.which, 2]),
															",\\ ",
															gsub("\\$", "", tab1.2[med.which, 3]),
															")$")

tab1.2 <- cbind(rownames(tab1.2), tab1.2[,1:2])

colnames(tab1.2)[1] <- "\\multicolumn{3}{l}{Men still alive in 1985}\\\\%"

tab1.2[-c(6:8, 12:16),] %>%
	sanitize.tex(firstcol = F) %>% xtable(align = "llrl") %>% print.xtable()

```
\end{table}

#

\begin{table}[H]
\centering
```{r}
tab1.3 <- readRDS(here::here('resources/', 'popchar3.tab1.rds'))

n.which <- which(grepl("\\$n", tab1.3[,"spread.which"]))
med.which <- which(grepl("median", tab1.3[,"spread.which"]) &
									 	!is.na(tab1.3[,3]))

tab1.3[med.which, 2] <- paste0("$(", gsub("\\$", "", tab1.3[med.which, 2]),
															",\\ ",
															gsub("\\$", "", tab1.3[med.which, 3]),
															")$")

tab1.3 <- cbind(rownames(tab1.3), tab1.3[,1:2])

colnames(tab1.3)[1] <- "\\multicolumn{3}{l}{Men still alive in 1985}\\\\%"

tab1.3[-c(6:8, 13:16),] %>% 
	sanitize.tex(firstcol = F) %>% xtable(align = "llrl") %>% print.xtable

```
\end{table}

# Cox PH Models

\linespread{1.25}

- Cumulative exposure to straight, soluble, and synthetic MWFs (lagged 20 years)
- Splined calendar year
- Splined year of hire
- Race
- Plant
- Risk sets indexed by age

#

## 1. Men still alive in 1973

\begin{table}[H]
\centering
```{r}
coef1.tab <- readRDS(here::here('reports/prostate cancer/resources/', "pr_mod1.tab.rds"))

colnames(coef1.tab)[5] <- "(95\\% CI)"
coef1.tab <- coef1.tab[,-c("events", "p")]
coef1.tab[,"n"] <- as.character(round(unlist(coef1.tab[,"n"]), 0))
coef1.tab[,"Covariate"] <- gsub(" 5$", "", unlist(coef1.tab[,"Covariate"]))

coef1.tab %>% 
	sanitize.tex %>% xtable %>% print.xtable
		
```
\end{table}

#

## 2. Men still alive in 1985

\begin{table}[H]
\centering
```{r}
coef2.tab <- readRDS(here::here('reports/prostate cancer/resources/', "pr_mod2.tab.rds"))

colnames(coef2.tab)[5] <- "(95\\% CI)"
coef2.tab <- coef2.tab[,-c("events", "p")]
coef2.tab[,"n"] <- as.character(round(unlist(coef2.tab[,"n"]), 0))
coef2.tab[,"Covariate"] <- gsub(" 5$", "", unlist(coef2.tab[,"Covariate"]))

coef2.tab %>%
	sanitize.tex %>% xtable %>% print.xtable

```
\end{table}

# 

## 3. Men still alive in 1985, weighted by probability of not being dead due to natural causes in 1985

\begin{table}[H]
\centering
```{r}
coef3.tab <- readRDS(here::here('reports/prostate cancer/resources/', "pr_mod3.tab.rds"))

colnames(coef3.tab)[5] <- "(95\\% CI)"
coef3.tab <- coef3.tab[,-c("events", "p")]
coef3.tab[,"n"] <- as.character(round(unlist(coef3.tab[,"n"]), 0))
coef3.tab[,"Covariate"] <- gsub(" 5$", "", unlist(coef3.tab[,"Covariate"]))

coef3.tab %>%
	sanitize.tex %>% xtable %>% print.xtable

```
\end{table}

#

## 4. Men still alive in 1985, weighted by the stabilized inverse probability with truncation

\begin{table}[H]
\centering
```{r}
coef4.tab <- readRDS(here::here('reports/prostate cancer/resources/', "pr_mod4.tab.rds"))

colnames(coef4.tab)[5] <- "(95\\% CI)"
coef4.tab <- coef4.tab[,-c("events", "p")]
coef4.tab[,"n"] <- as.character(round(unlist(coef4.tab[,"n"]), 0))
coef4.tab[,"Covariate"] <- gsub(" 5$", "", unlist(coef4.tab[,"Covariate"]))

coef4.tab %>% sanitize.tex %>% xtable %>% print.xtable

```
\end{table}

## 5. Men still alive in 1985, weighted by the stabilized inverse probability with truncation

\begin{table}[H]
\centering
```{r}
coef4.tab <- readRDS(here::here('reports/prostate cancer/resources/', "pr_mod4.tab.rds"))

colnames(coef4.tab)[5] <- "(95\\% CI)"
coef4.tab <- coef4.tab[,-c("events", "p")]
coef4.tab[,"n"] <- as.character(round(unlist(coef4.tab[,"n"]), 0))
coef4.tab[,"Covariate"] <- gsub(" 5$", "", unlist(coef4.tab[,"Covariate"]))

coef4.tab %>% sanitize.tex %>% xtable %>% print.xtable

```
\end{table}




# Results by MWF type

\begin{table}[H]
\centering
```{r}
str.tab <- cbind(
	coef1.tab[grep("str", zoo::na.locf(coef1.tab[,Covariate])), 2],
	NA,
	coef1.tab[grep("str", zoo::na.locf(coef1.tab[,Covariate])),-c(1, 2, 5)],
	NA,
	coef2.tab[grep("str", zoo::na.locf(coef2.tab[,Covariate])), c(3, 4)],
	NA,
	coef3.tab[grep("str", zoo::na.locf(coef3.tab[,Covariate])), c(3, 4)],
	NA,
	coef4.tab[grep("str", zoo::na.locf(coef4.tab[,Covariate])), c(3, 4)])

colnames(str.tab) <- sapply(1:ncol(str.tab), function (x) {paste(rep("%", x), collapse = "")})

str.tab %>%
	sanitize.tex %>% xtable(align = "llcccccccccccc") %>% print.xtable(
	add.to.row = list(
		pos = list(0),
		command = paste0(
			"Cumulative straight & &",
			"\\multicolumn{2}{c}{1973} & &",
			"\\multicolumn{2}{c}{1985} & &",
			"\\multicolumn{2}{c}{1985, weighted} & &",
			"\\multicolumn{2}{c}{1985, s. weight}",
			"\\\\\n",
			"\\cline{3-4}",
			"\\cline{6-7}",
			"\\cline{9-10}",
			"\\cline{12-13}",
			"\n",
			paste0(rep("&& n & HR", 4), collapse = ""),
			"\\\\")
	))

```
\end{table}

\begin{table}[H]
\centering
```{r}
sol.tab <- cbind(
	coef1.tab[grep("sol", zoo::na.locf(coef1.tab[,Covariate])), 2],
	NA,
	coef1.tab[grep("sol", zoo::na.locf(coef1.tab[,Covariate])),-c(1, 2, 5)],
	NA,
	coef2.tab[grep("sol", zoo::na.locf(coef2.tab[,Covariate])), c(3, 4)],
	NA,
	coef3.tab[grep("sol", zoo::na.locf(coef3.tab[,Covariate])), c(3, 4)],
	NA,
	coef4.tab[grep("sol", zoo::na.locf(coef4.tab[,Covariate])), c(3, 4)])

colnames(sol.tab) <- sapply(1:ncol(sol.tab), function (x) {paste(rep("%", x), collapse = "")})

sol.tab %>%
	sanitize.tex %>% xtable(align = "llcccccccccccc") %>% print.xtable(
		add.to.row = list(
			pos = list(0),
			command = paste0(
				"Cumulative soluble & &",
				"\\multicolumn{2}{c}{1973} & &",
				"\\multicolumn{2}{c}{1985} & &",
				"\\multicolumn{2}{c}{1985, weighted} & &",
				"\\multicolumn{2}{c}{1985, s. weight}",
				"\\\\\n",
				"\\cline{3-4}",
				"\\cline{6-7}",
				"\\cline{9-10}",
				"\\cline{12-13}",
				"\n",
				paste0(rep("&& n & HR", 4), collapse = ""),
				"\\\\")
		))

```
\end{table}

\begin{table}[H]
\centering
```{r}
syn.tab <- cbind(
	coef1.tab[grep("syn", zoo::na.locf(coef1.tab[,Covariate])), 2],
	NA,
	coef1.tab[grep("syn", zoo::na.locf(coef1.tab[,Covariate])),-c(1, 2, 5)],
	NA,
	coef2.tab[grep("syn", zoo::na.locf(coef2.tab[,Covariate])), c(3, 4)],
	NA,
	coef3.tab[grep("syn", zoo::na.locf(coef3.tab[,Covariate])), c(3, 4)],
	NA,
	coef4.tab[grep("syn", zoo::na.locf(coef4.tab[,Covariate])), c(3, 4)])

colnames(syn.tab) <- sapply(1:ncol(syn.tab), function (x) {paste(rep("%", x), collapse = "")})

syn.tab %>% sanitize.tex %>% xtable(align = "llcccccccccccc") %>% print.xtable(
						 add.to.row = list(
						 	pos = list(0),
						 	command = paste0(
						 		"Cumulative synthetic & &",
						 		"\\multicolumn{2}{c}{1973} & &",
						 		"\\multicolumn{2}{c}{1985} & &",
						 		"\\multicolumn{2}{c}{1985, weighted} & &",
						 		"\\multicolumn{2}{c}{1985, s. weight}",
						 		"\\\\\n",
						 		"\\cline{3-4}",
						 		"\\cline{6-7}",
						 		"\\cline{9-10}",
						 		"\\cline{12-13}",
						 		"\n",
						 		paste0(rep("&& n & HR", 4), collapse = ""),
						 		"\\\\")
						 ))

```
\end{table}


___