---
title: "UAW-GM Cohort Study"
subtitle: "Predicting survival to 1985"
# fontsize: 10pt
date: \today
# mainfont: Arial
# Windows only:
# mainfontoptions: "Extension = .ttf, UprightFont = \\*, BoldFont = \\*bd, ItalicFont = \\*i, BoldItalicFont = \\*bi"
output:
  # pdf_document:
  #   latex_engine: lualatex
  #   includes:
  #     in_header: "~/HeadRs/StatHead.sty"
  #   number_sections: false
  #   toc: false
  #   keep_tex: true
  # word_document:
  #   reference_docx: "~/HeadRs/style-guide-arial-11.dotx"
  beamer_presentation:
    toc: false
    slide_level: 1
    includes:
      in_header: "~/HeadRs/BeamerHead.sty"
# bibliography: ./../../references/gm.bib
# csl: "C:/Users/kevchen/Documents/HeadRs/AMA.csl"
# geometry: margin=2.54cm
---

```{r setup, include=F}
# setwd('gm')
knitr::opts_chunk$set(
	echo = F,
	# eval = F,
	warning = F,
	message = F,
	fig.align = "center",
	fig.pos = "H",
	results = "asis")

library(here); library(knitr); library(tidyverse); library(pander); library(xtable)
library(data.table); library(lubridate); library(Hmisc);
source("~/HeadRs/00-my-theme.R")

table.engine <- "xtable"
# table.engine <- "pander"

weights.year.min <- 1941

```

# Population

\linespread{1.5}

- Restricted to those:
	- Still alive in `r weights.year.min`
	- Hired in or after 1938, but no later than 1982
	- Missing no more than half of their work record
- Individuals contributed person-time from three years after hire or `r weights.year.min` (whichever came first) to death or loss to follow-up
- Individuals were considered lost to follow-up upon reaching the oldest observed age at death (106.56 years)
- N = 36 549, (698 394 person-years)
- Deaths due to natural causes by end of 1984: 4 405 (11.4\%)

# ICD codes for natural causes of death

\linespread{1.5}

- ICD-9: all codes codes in [001, 799]
	- Excludes the categories labeled as “Injury and poisoning” and “external causes of injury and supplemental classification.”
- ICD-10: all codes, except those with prefix S, T, V, W, X, or Y.

# Estimating survival in each year using Super Learner

\linespread{1.25}

- We use the Super Leaner [`tlverse/sl3`](https://github.com/tlverse/sl3) to estimate the probability dying due to natural causes by the end of each year of follow-up, conditional on covariates and survival prior to that year
- The Super Learner library included:
	- Covariate combination-specific mean `mean`
	- Pooled logistic regression `glm`
	- PLR with reglarization by lasso or elasticnet `glmnet`
	- Generalized additive model `gam`
	- Random forests `ranger`
	- Extreme gradient boosting: additive decision tree ensembles `xgboost`

# Covariates

\linespread{1.25}

- Duration of employment
- Calendar year
- Age
- Race
- Plant
- Sex
- Cumulative time spent off
- Year of hire
- Cumulative exposure to straight, soluble, and synthetic MWFs
- Employment status

<!-- # Model 1 results -->

<!-- ```{r} -->
<!-- glm.tab <- readRDS(here::here(paste0("reports/survival to 1985/resources/FU from ", weights.year.min), "small.glm.tab.rds"))  -->
<!-- glm.tab[, `:=`(Covariate = { -->
<!-- 	tmp <- Covariate -->
<!-- 	tmp[duplicated(tmp)] <- "" -->
<!-- 	tmp -->
<!-- })] -->
<!-- ``` -->

<!-- \begin{table}[H] -->
<!-- \begin{adjustbox}{scale=0.75} -->
<!-- \centering -->
<!-- ```{r} -->
<!-- xtable(glm.tab[1:20,-"SE"]) -->
<!-- ``` -->
<!-- \end{adjustbox} -->
<!-- \end{table} -->

<!-- # -->

<!-- ## Model 1 results (continued) -->

<!-- \vspace{2.5pt} -->
<!-- \begin{table}[H] -->
<!-- \begin{adjustbox}{scale=0.75} -->
<!-- \centering -->
<!-- ```{r} -->
<!-- xtable(glm.tab[21:nrow(glm.tab),-"SE"]) -->
<!-- ``` -->
<!-- \end{adjustbox} -->
<!-- \end{table} -->

<!-- # Model 2 results -->

<!-- ```{r} -->
<!-- gam.tab <- readRDS(here::here("reports/survival to 1985/resources", "small.gam.tab.rds")) -->
<!-- gam.tab[, `:=`(Covariate = { -->
<!-- 	tmp <- Covariate -->
<!-- 	tmp[duplicated(tmp)] <- "" -->
<!-- 	tmp -->
<!-- })] -->
<!-- ``` -->

<!-- \begin{table}[H] -->
<!-- \begin{adjustbox}{scale=0.75} -->
<!-- \centering -->
<!-- ```{r} -->
<!-- xtable(gam.tab[1:17,-"SE"]) -->
<!-- ``` -->
<!-- \end{adjustbox} -->
<!-- \end{table} -->

<!-- # -->

<!-- ## Model 2 results (continued) -->

<!-- \vspace{\baselineskip} -->
<!-- \begin{table}[H] -->
<!-- \begin{adjustbox}{scale=0.75} -->
<!-- \centering -->
<!-- ```{r} -->
<!-- xtable(gam.tab[18:nrow(gam.tab),-"SE"]) -->
<!-- ``` -->
<!-- \end{adjustbox} -->
<!-- \end{table} -->
<!-- \vspace{\baselineskip} -->
<!-- \footnotesize Splined terms not shown. -->

# Calculating cumulative survival

\linespread{1.25}

1. Extract the fitted probabilities $\hat p_{ti}$ from the learner(s)
2. For each individual, with obserations ordered by time $t$, take the cumulative product $\prod_t(1 - \hat p_{ti})$
3. The cumulative product in each row represents the probability of survival (for natural cause mortality) to the end of that row's year

# Average survival probabilities by age group

```{r}
for (covariate in c("Age", "Employment status", "Race", "Sex", "Year of hire",
										paste("Cumulative", c("straight", "soluble", "synthetic")))) {
	tmp <- readRDS(
				 	file = here::here("resources", paste0("survival_by-", gsub(" ", "-" , covariate), ".rds"))
				 )
	
	if (sum(grepl("\\]", unlist(tmp[,1]))) > 0) {
	tmp[,(covariate) := paste0("$", get(covariate), "$")]
		}
	
	assign(paste0(gsub(" ", "_" , tolower(covariate)), ".prob.tab"),
				 tmp,
				 envir = .GlobalEnv)
}

bytimevarying.tab <- rbindlist(list(
	age.prob.tab,
	employment_status.prob.tab), idcol = "Covariate", use.names = F)

names(bytimevarying.tab)[2] <- "level"

bytimevarying.tab[,`:=`(Covariate = {
	tmp <- factor(Covariate, 1:2, c("Age", "Employment status"))
	tmp <- levels(tmp)[as.numeric(tmp)]
	tmp[duplicated(tmp)] <- ""
	tmp
	},
	level = gsub(",", ", ", level))]

```

\begin{table}[H]
\begin{adjustbox}{scale=0.75}
\centering
```{r}
xtable(age.prob.tab)
```
\end{adjustbox}
\end{table}

# Average survival probabilities by other covariates

```{r}
invisible(
sapply(paste0(c("race", "sex", "year_of_hire", "employment_status"), ".prob.tab"),
			 function(tab) {
			 	cat(paste0("\\begin{table}[H]\n",
					 "\\begin{adjustbox}{scale=0.75}\n",
					 # "\\centering",
					 NULL
					 ))
			 	print.xtable(xtable(get(tab)))
			 	cat(paste0("\\end{adjustbox}",
			 						 "\\end{table}\n\n"))
			 }))
```


# Average survival probabilities by MWF exposure

\begin{table}[H]
\begin{adjustbox}{scale=0.75}
\centering
```{r}
bymwf.tab <- rbindlist(list(
	cumulative_straight.prob.tab, cumulative_soluble.prob.tab, cumulative_synthetic.prob.tab), idcol = "Cumulative exposure", use.names = F)

names(bymwf.tab)[2] <- "level"

bymwf.tab[,`:=`(`Cumulative exposure` = {
	tmp <- factor(`Cumulative exposure`, 1:3, c("Straight", "Soluble", "Synthetic"))
	tmp <- levels(tmp)[as.numeric(tmp)]
	# tmp[duplicated(tmp)] <- NA
	tmp
	},
	level = gsub(",", ", ", level))]

bymwf.tab <- rbindlist(lapply(unique(bymwf.tab$`Cumulative exposure`), function (x) {
	rbindlist(list(
		data.table(level = paste0("\\multicolumn{2}{l}{", x, "}\\\\\n%")),
		bymwf.tab[`Cumulative exposure` == x, -1, with = F]),
		fill = T)
}))

names(bymwf.tab)[1] <- "Cumulative exposure"

bymwf.tab %>% xtable(align = rep("r", ncol(.) + 1)) %>% print.xtable()

```
\end{adjustbox}
\end{table}

<!-- \begin{table}[H] -->
<!-- \begin{adjustbox}{scale=0.75} -->
<!-- \centering -->
<!-- ```{r} -->
<!-- xtable(bymwf.tab) -->
<!-- ``` -->
<!-- \end{adjustbox} -->
<!-- \end{table} -->

# ROC Curve

\includegraphics[scale=0.85]{`r here::here("reports/survival to 1985/resources", "sl.roc.pdf")`}

\footnotesize

Outcome: Natural cause mortality status in 1984

An individual's probability of not dying due to natural causes was calculated as: $\prod_t(1 - \hat p_t)$ where $\hat p_t$ is the predicted probability of death due to natural causes for the $t^\text{th}$ year of follow-up.

# ROC Thresholds

\begin{table}[H]
\begin{adjustbox}{scale=0.75}
\centering
```{r}
roc.threshold <- readRDS(
	file = here::here("resources", "roc.threshold.rds"))

names(roc.threshold) <- gsub("threshold\\.", "", names(roc.threshold))

roc.threshold %>% xtable(digits = 4) %>% print
```
\end{adjustbox}
\end{table}